---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(getrect)
```

# getrect

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
<!-- badges: end -->

`{getrect}` is a solution to the problem of partitioning a matrix of values
into 


## What's in the box

* `partition_matrix()` partition a matrix into rectangles 
* `plot_rects()` Plot the result of `partition_matrix()`
* `max_rect_under_histogram()` calculates the maximum area rectangle under
  a sequence of values representing bar heights of a histogram
* `plot_hist_max_rect()` plot the bar heights and best rectangle

## Future ideas

* There are faster algorithms for 'max area under a histogram' than what I've 
  included here.
* When only rectangles with area = 1 remain, could short-circuit the iteration
  and add them all in bulk.

## Installation

You can install from [GitHub](https://github.com/coolbutuseless/getrect) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/getrect')
```


# Problem: Partition matrix into rectangular regions

Partition a matrix into rectangular regions where:

* Each region contains only a single unique value
* An effort is made to minimise the total number of rectangles 


# Solution:

* For each unique value: 
    * Greedily find the biggest single-valued rectangle at each step



## Key algorithm: Maximum rectangular area under a histogram

Given a sequence of bar heights (representing a histogram), what is the 
maximum sized rectangle which can be inscribed?

```{r example}
library(getrect)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Define bar heights
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
hts <- c(4, 8, 3, 1, 3, 3, 0, 1)
plot_hist_max_rect(hts)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Calculate maximum sized rectangle
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
rect <- max_rect_under_histogram(hts)
plot_hist_max_rect(hts, rect)
```





# Step-by-step example on example matrix

```{r}
mat <- matrix(
  c(1,1,1,2,3,3,3,
    1,1,2,2,2,3,3,
    1,2,2,2,2,2,3,
    2,2,2,2,2,2,2), 
  nrow = 4, ncol = 7, byrow = TRUE
)
mat
```




## Consider a single target value

Consider just the value `2` in the matrix.  

Create a working matrix `s` with a `1` where the target value exists, and `0`
elsewhere.

```{r}
target <- 2
s <- mat
s[] <- mat == 2
s
```

## Calculate the vertical runs

Calculate the vertical runs of `1` at every row based upon the rows above

```{r}
s <- apply(s, 2, \(x) ave(x, cumsum(x == 0), FUN = cumsum))
s
```

## Consider each row in isolation, and calculate the *max rect under the histogram*

```{r}
row_max_rects <- apply(s, 1, \(x) as.data.frame(max_rect_under_histogram(x)))
row_max_rects <- do.call(rbind, row_max_rects)
row_max_rects
```

## Pick the best row (max rect area)

```{r}
idx <- which.max(row_max_rects$area)
best_rect <- row_max_rects[idx, ]

# Calculate the y extents of the rectangle
best_rect$ymax <- idx
best_rect$ymin <- idx - best_rect$h + 1
best_rect$target <- target
```

## Mark part of matrix as done

```{r}
plot_rects(best_rect, mat)
```


## Repeat until no unmarked areas remain

```{r echo = FALSE}
rects <- partition_matrix(mat)
plot_rects(rects, mat)
```

# Larger example

```{r}
set.seed(2)
w <- 15
h <- 15
mat <- matrix(sample(c(1, 2), w*h, replace = TRUE, prob = c(4, 1)), h, w)
mat

# Partition
rects <- partition_matrix(mat)
plot_rects(rects)
```



# Extension: prefer areas with an aspect ratio close to 1

Same matrix as above, but penalise the selection of rectangles with
high aspect ratio by setting `ar_penalty`.

```{r}
rects <- partition_matrix(mat, ar_penalty = 0.9)
plot_rects(rects)
```



